\Huge{\textbf{Using Blockchain to Create a Decentralised Security Model for Distributed Systems}}

### 2021-04-16 paper/paper.tex
YEET WE MIGHT BE DONE

%%%%%%%%%%%%%%%%%%
% ABSTRACT
%%%%%%%%%%%%%%%%%%
\begin{abstract}
Network security has, and always will be one of the greatest concerns for universities and other educational institutions. The 2020 cyberattacks throughout the COVID-19 pandemic highlighted both the importance of cybersecurity in UK universities, but also the lack of modern security infrastructure deployed across campuses. This literature aims to provide a method for universities to share information on potential, or ongoing cyberattacks to other universities via a trustworthy channel, capable of operating independently from other security measures. This is achieved by creating middleware capable of running on existing devices on each campus that uses blockchain to establish a decentralised trust mechanism. We explore a number of methods to achieve this including examining existing technologies as well as designing and implementing our own. A thorough analysis of the provided implementation is provided and we conclude by suggesting a number of future improvements that could be made to improve both the efficiency and effectiveness of such a system.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%
% INTRODUCTION
%%%%%%%%%%%%%%%%%%
\chapter{Introduction}
\section{COVID-19 and Cyberattacks}
In the summer of 2020 during the midst of the COVID-19 pandemic, universities and research institutions worldwide were working hard to understand the structure of the virus and develop a vaccine in an attempt to return to normality. However, whereas some countries were making fast progress in understanding the virus, others were falling behind, and the virus began to put a strain on healthcare, and increasing critique on governments. In order to keep up with the nations at the forefront of vaccine development, nations turned to state-sponsored cyberattacks in order to both hinder nations, and also obtain research and information about other countries' vaccine efforts. One such example was the threat group 'Cozy Bear', formally known as \acrfull{APT} 29. \acrshort{APT}29 used a number of tools to target various organisations involved in COVID-19 vaccine development in Canada, the United States and the United Kingdom. The \acrfull{NCSC} believe that the intention was highly likely stealing information and intellectual property relating to the vaccine \cite{APT29}.

@@

University attacks became a frequent headline in the UK as universities suffered attacks from different threat actors. A number of threat actors launched attacks against multiple universities in the hope to find a vulnerability in at least one. One such attack was aimed at both Newcastle University and Northumbria University, two universities in extremely close proximity \cite{newhack,norhack}. The attack crippled both Newcastle and Northumbria Universities, however the attackers only managed to exfiltrate data from Newcastle University. Why was the attack successful on both occasions? Why wasn't knowledge of the attack shared?

@@

@@

\begin{center}
	\begin{figure}[!htb]
		\includegraphics[width=\textwidth,keepaspectratio]{TanenbaumDistributed}
		\caption{A Distributed System visualised as \gls{middleware} \cite{tanenbaumdist}} 
		\label{fig:middleware}
	\end{figure}
\end{center}

@@

\section{Aim} \label{aim}
The original aim for this project was to design and create a decentralised firewall that could communicate knowledge of cyberattacks aimed at universities in real-time, allowing other universities to protect themselves from the same attacks. This system would be distributed, and hence must conform to the previous description of a distributed system in section \ref{distributed}. Following an extensive amount of background reading, there appeared to be no existing implementation or design of such a system which inspired me to alter my aim and instead focus entirely on designing a protocol and implementing a \gls{stub} to demonstrate the protocol's effectiveness. This project will therefore not be implementing a firewall, but instead a system to coordinate firewalls. Further research determined that \gls{blockchain} was the best choice for the underlying structure for such a protocol, and so this final change shaped the current aim for this project: \textbf{Using Blockchain to Create a Decentralised Security Model for Distributed Systems}.

@@

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth,keepaspectratio]{rpc}
\caption{A breakdown of an \acrshort{RPC} \cite{tanenbaumdist}} 
\label{fig:rpc}
\end{figure}

@@

\section{Distributed Security}
The primary reference used for distributed security was Rivets and Lampson's \say{SDSI - A Simple Distributed Security Infrastructure} \cite{sdsi}. This publication provides an in-depth explanation of how a public-key infrastructure can be used in conjunction with access control lists to create a distributed security infrastructure. The majority of the literature within this publication is focussed on creating and issuing certificates, something that is not relevant for this project, however Rivets and Lampson did provide clear requirements over the data structures within such a system. Rivets and Lampson implement a message system similar to that of Tanenbaum and van Steen's in section \ref{dist}. The message system proposed by Rivets and Lampson contains only a type and dictionary of attributes.

@@

\section{Fault Tolerance}
The primary reference for fault tolerance was Guerraoui and Schiper's \say{Fault-Tolerance by Replication in Distributed Systems}\cite{faulttol}. This literature details how replication can be used to provide fault tolerance in a distributed system in addition to ensuring consistency is maintained.  A number of backup techniques are discussed however the technique that is best suited for application is primary backup replication. Primary backup replication consists of a client invoking an operation which is the applied to the primary data, and then cascaded to a number of additional backups.

@@

The type field holds an enum which is called MessageType and contains three values: advertisement, consensus and rule. Similarly the subtype field holds an enum that contains two values: broadcast and acknowledgement. Broadcast messages are used to advertise a new device, request consensus and distribute new a firewall transaction, whereas acknowledgements are used to provide a response, such as acknowledging the new device and providing consensus for a proposed transaction. As the system is decentralised, messages are not sent directly to remote hosts as there may be devices on the network that another device is unaware of, which would result in unsynchronised \gls{blockchain}s. The use of a hop count allows us to limit the number of times a message can be sent around the network, essential to prevent older messages from clogging up the system. The last three fields contain addresses of devices. It may appear unnecessary to provide these values as the messages are wrapped in an \acrshort{IP} packet which already contains these values. These are necessary as when messages are relayed, the \acrshort{IP} packet contains the address of the most recent sender, overwriting the original information. The source address contains the address of the origin device, the target address contains the address of the final destination, and the next address contains the address of the next device to propagate the message to.

\subsubsection{Advertisement Messages}
The advertisement message is used to advertise the presence of a new device within the distributed system. This message contains no additional fields than those described in table \ref{tab:common}. When the advertisement message subtype is broadcast, the source address is set to the new device's \acrshort{IP} address. At least one host must be known to advertise on the network, and this host's address will be placed into the next address field. The target address field is not used for broadcasts as we have no specific target that we wish to send our message to. Once a node receives an advertisement broadcast, it will check whether that host is already known, if so, the message is ignored, and propagated, otherwise if the host is not yet known, the node first adds the new device to its list of known hosts, and then propagates the message. When the acknowledgement is sent, the target address is set to the address of the new device, which ensures that any other nodes which may have advertised do not interpret the acknowledgement as regarding their broadcast. Once the acknowledgement arrives at the new device, the node adds the host which sent the ack, but does not propagate the message.

\subsubsection{Consensus Messages}
The consensus message is used when a client or firewall submits a new firewall rule to the \gls{stub} via \acrshort{IPC}. This message contains one additional field which holds the hash of the last block. When a node proposes a new firewall rule, it must first submit the hash of it's last block, which is sent to all known hosts. On receipt of a consensus broadcast, the node calculates the hash of the last block on it's chain, and if it matches the hash sent in the message it sends an ack, and appends the origin node's address to a list of pending rules. The node then forwards the message to all known hosts. After broadcasting the message, the origin host resets a counter, which is incremented every time it receives an ack from one of it's known hosts. If this counter reaches atleast half of it's known hosts in some bounded time, it is deemed to have obtained consensus and may now distributed the new firewall rule.

@@

\subsection{IPC}
There are four different types of \acrshort{IPC} message: rule, enable, disable, shutdown. The rule type is intended to be sent by a firewall, and contains a new firewall rule to be submitted to the distributed system. The enable and disable message types simply control the status of the stub, if it is believed that a host on the network may be malfunctioning (i.e. a firewall is sending bogus rules), it can be disabled and then re-enabled once the fault is fixed. Finally the shutdown message is used to properly terminate the \gls{stub}. Upon receiving the shutdown message, all sockets will be closed, the \acrshort{IPC} tunnel will be destroyed and the application will wait for all threads to terminate before exiting.
To ensure messages can be parsed correctly, all of these message types follow the same structure and simply do not use the redundant fields.

@@

\subsection{Advertisement}
The advertisement message is sent at startup and is mutually exclusive from the other messages, as it is executed automatically and does not affect the consensus or rule transaction stages. The advertisement message is sent when the framework is executed. The framework will send an advertisement broadcast to all known hosts, each of which will relay the message to each of their known hosts, provided that the hop count has not yet been exceeded. The known host will then forward the message to all of its known hosts, propagating the message throughout the network. Upon receiving an advertisement message, the host checks if the host is known, if not, it is added, and an acknowledgement is returned.

@@

\subsection{Consensus-Rule}
The consensus and rule messages are sent when a client or firewall send an \acrshort{IPC} message to the local \gls{stub}. Upon receiving the rule, the \gls{stub} will marshal the parameters into a network message, which is then sent to all known hosts. When a node receives the consensus message, it will compare the hash in the message to the hash of the previous block on its chain, if the hashes match, the block is considered valid and an acknowledgement is returned. The broadcast is then forwarded to all of the hosts known by that node. When the origin node receives an acknowledgement, it will check if that host is known. If it is not, then it will not increment the ack count, as it would interfere with the consensus calculation, however it will still receive the rule, if consensus is achieved. The origin node will then wait for some timeout, after which it will test if a sufficient number of acknowledgements have been received in comparison to the number of known hosts. If this test passes (e.g. at least half of the known hosts must acknowledge), then a rule message is sent. No acknowledgements are sent upon receipt of a rule message, but nodes will continue to propagate the message until the hop limit is reached.

@@

\section{IPC}
In order to achieve cross-platform \acrshort{IPC}, different approaches were required for \acrfull{POSIX} compliant operating systems and Windows\cite{windows} operating systems. Each \acrshort{OS} type offers a number of \acrshort{IPC} methods, which are briefly discussed the following sections.

The \gls{stub} provides the following abstractions for interfacing with native \acrshort{IPC}. The implementation methods are discussed in their relative sections below.\\\\

@@

\subsection{POSIX} \label{posixipc}
\acrshort{POSIX} systems provide four interfaces for \acrshort{IPC}: pipes, FIFO, message queues and sockets. Although pipes do allow processes to communicate, the primary purpose of pipes is to obtain the output of one command and pass this output into a second command, which is not suitable for our application. FIFO and message queues operate very similarly, whereby each method creates some file in the filesystem which can then be referenced by any process. FIFO is the predecessor to message queues, and hence different operating systems have varying implementations for FIFO, however as message queues are a \acrshort{POSIX} standard, all implementations provide the same interface, greatly simplifying the processes of making this project cross compatible. For this reason, message queues were used for \acrshort{POSIX} \acrshort{IPC}. The final method is via UNIX sockets. The socket interface is primarily designed for networked communication, but also supports \acrshort{IPC}. Unfortunately, as sockets are intended for use in networks, some implementations still use the \acrshort{NIC} to communicate even locally, which slows down communication, and adds additional points of failure.

@@

\section{Sockets}
Network sockets are the primary method allowing processes to communicate over a network, however similarly to \acrshort{IPC}, \acrshort{POSIX} and Windows system provide slightly different socket interfaces. A numeric port is bound to the socket which enables the network stack to communicate with the socket, and hence the process. The main difference is that Windows requires the initialisation of the WinSock\cite{winsock} library. Thankfully other than the initialisation of WinSock, the rest of the interface provided by WinSock is very similar to the basis of \acrshort{POSIX} sockets, which are known as Berkeley sockets. This similarity means that only small modification need to be made to the source code. After creating the socket, it is then bound to the numeric port. The default ports used for this project are 8070 and 8071.

@@

\section{Multithreading}
In order for the \gls{stub} to be capable of consecutively send and receive messages over both the network, and via \acrshort{IPC}, the framework had to be multithreaded. Unsurprisingly, \acrshort{POSIX} and Windows use different thread models, which means there is no common threading interface. Thankfully, the \acrshort{POSIX} thread model has been ported to Windows in the form of the \acrshort{POSIX} Threads for Windows project\cite{pthread}. This allows the source code to be written identically for all operating systems, the only difference being that Windows will be linked with the necessary Dynamic-Link Library (DLL).

\section{Blockchain}
Once the previous \acrshort{API}s had been created, the \gls{blockchain} aspect of the project could be implemented. As the structure of these blocks have already been discussed in section \ref{firewallblock}, this section will focus on how the functionality of this was implemented, primarily the hashing of blocks.

\subsection{Hash Algorithms} 
The first decision that had to be made was which hash algorithm to implement. A wide range of hash algorithms are available, however the most popular three are Secure Hash Algorithm (SHA), MD5 and BLAKE. The SHA hash suite boasts a number of algorithms which are classed into four generations: SHA-0, SHA-1, SHA-2 and SHA-3. The SHA-0 and SHA-1 algorithms are the least secure, and are now deemed insecure as collisions been found, which would enable a malicious block to contain the hash of a valid block. At the time of writing this, there are no known collisions for the SHA-2 or SHA-3 families. The MD5 hash algorithm is also insecure for the same reason as SHA-1 and is hence not appropriate for this project. The final algorithm family is BLAKE, which consists of three families: BLAKE, BLAKE2 and BLAKE3. There are no known collisions within any of the BLAKE hash families, however the BLAKE algorithm has not been tested for \gls{blockchain} applications, whereas the SHA family has. For this reason, this project will utilise the same hash algorithm as Bitcoin \cite{bitcoin}, SHA-256.

\begin{table}[H]
\centering
\begin{tabular}{ |c|c|c|c|c| } 
\hline
\multicolumn{2}{|c|}{Hash Algorithm} & Output Size & Collisions found? & Used By \\
\hline
\multicolumn{2}{|c|}{MD5} & 128 & Yes & None \\
\hline
\multicolumn{2}{|c|}{SHA-0} & 160 & Yes & Nonimprovee \\
\hline
\multicolumn{2}{|c|}{SHA-1} & 160 & Yes & None \\
\hline
\multirow{6}*{SHA-2} & SHA-224 & 224 & No & None \\
& SHA-256 & 256 & No & Bitcoin\cite{bitcoin}, Bitcoin Cash \\
& SHA-384 & 384 & No & None \\
& SHA-512 & 512 & No & None \\
& SHA-512/224 & 224 & No & None \\
& SHA-512/256 & 256 & No & None \\
\hline
\multirow{4}*{SHA-3} & SHA3-224 & 224 & No & None \\
& SHA3-256 & 256 & No & None \\
& SHA3-384 & 384 & No & None \\
& SHA3-512 & 512 & No & None \\
\hline
\multirow{4}*{BLAKE} & BLAKE-224 & 224 & No & None \\
& BLAKE-256 & 256 & No & None \\
& BLAKE-384 & 384 & No & None \\
& BLAKE-512 & 512 & No & None \\
\hline
\multirow{2}*{BLAKE2} & BLAKE2s & 256 & No & Nano\cite{nano} \\
& BLAKE2b & 512 & No & None \\
\hline
\multicolumn{2}{|c|}{BLAKE3} & Variable & No & None \\
\hline
\end{tabular}
\caption{Comparison of Hash Algorithms}
\label{tab:hashalgo}
\end{table}

\subsection{Hash Implementation}
When implementing the desired hash algorithm, there were to primary options, to implement the algorithm from scratch, or to use an existing library. If the algorithm was to be implemented from scratch, then the resulting application would requires less dependencies and would overall result in a more self-contained binary. Using a library would result in one additional dependency, however the vast number of algorithms provided by most libraries would allow the entire \gls{blockchain} and protocol to be easily updated should a more secure algorithm be preferable. Furthermore, most libraries are thoroughly tested and well engineered to guarantee the most efficient operation and make sure the algorithms can handle every edge case. For this it was decided that an external library would be implemented, and a breakdown of the considered libraries is provided below \cite{hashlibcomp}.

@@

From the previous table, it is apparent that libgcrypt supports the most algorithms from the comparison given in \cite{hashlibcomp}, however for cross compatibility, cryptlib boasts 37 supported operating systems. OpenSSL's libcrypto sits in the middle of the two, supporting nine hash algorithm families and 28 different operating systems. Additionally, libcrypto is has been the default cryptography library distributed with most Linux, BSD and Windows operating systems, meaning it has received a large amount of support, and faced tight scrutiny by the likes of Microsoft. This made libcrypto the optimum library for use in this project.

\section{Fault Tolerance}
In regards to fault tolerance, this project was extremely limited, as it ran as a high-level application on general operating systems, meaning the operating systems themselves provided no guarantees of  performance or reliability. With regards to fault tolerance during operation, the only reasonable measure was to ensure no malformed data could be propagated throughout the system. Such malicious data could cause the local \gls{stub}, or even worse, remote \gls{stub}s to crash, rendering the entire system non-functional whilst the malicious data is removed from each system. This measure was achieved by using the fail-early principle in all functions within the system. When each function is executed, specific checks take place, such as ensuring buffers are large enough to store specified data, and that the data provided is valid, and of the required type. Many of these checks are provided by the standard library within the C language, such as verifying the string representation of an \acrshort{IP} address is  correctly formatted when parsing to a binary representation of the network address. The vast majority of these functions return an integer to represent the whether the function exited normally, or if it exited with an error, which was replicated in many of the functions provided by this application. Returning the exit status enables any calling functions to know that the function did not exit normally, which means the output of that function will contain incorrect data, and so the path of execution can no longer follow the normal path, but must instead fail early, and return to the top-level calling function.

In addition to the fail-early principle, the each node maintains a log of all complete transactions that have taken place on that node. Once a node receives a either an advertisement, or a new rule, it is logged immediately. This means that should the operating system fail, or the system suffers from an unrecoverable fault, the operating system can immediately restart the system (which can run as a daemon) which will then load all known hosts into memory, and rebuild the chain from the transaction log.

\chapter{Testing} \label{test}
Testing the system required tests to be categorised into two types: local and network. Local tests consist of any functions which take place on the local node, such as hashing of blocks and controlling network sockets. These local tests could be easily tested using unit tests. Network tests consist of protocol tests that can only be tested over a physical network. Network tests are much harder to strategically test, which means unit tests would require careful synchronisation, such that when one node was testing it's send feature, another node would need to be ready to accept a message, for this reason, networks tests did not use unit tests.

>>>>

%%%%%%%%%%%%%%%%%%
% ABSTRACT
%%%%%%%%%%%%%%%%%%
\begin{abstract}
Network security has, and always will be one of the greatest concerns for universities and other educational institutions. The 2020 cyberattacks throughout the COVID-19 pandemic highlighted both the importance of cybersecurity in UK universities, but also the lack of modern security infrastructure deployed across campuses. This literature aims to provide a method for universities to share information on potential, or ongoing cyberattacks to other universities via a trustworthy channel, capable of operating independently from other security measures. This is achieved by creating \gls{middleware} capable of running on existing devices on each campus that uses \gls{blockchain} to establish a decentralised trust mechanism. We explore a number of methods to achieve this including examining existing technologies as well as designing and implementing our own. A thorough analysis of our implementation is provided and we conclude by suggesting a number of future improvements that could be made to improve both the efficiency and effectiveness of such a system.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%
% INTRODUCTION
%%%%%%%%%%%%%%%%%%
\chapter{Introduction}
\section{COVID-19 and Cyberattacks}
In the summer of 2020, during the midst of the COVID-19 pandemic, universities and research institutions worldwide were working hard to understand the structure of the virus and develop a vaccine in an attempt to return to normality. However, whereas some countries were making fast progress in understanding the virus, others were falling behind, and the virus began to put a strain on healthcare, and increasing critique on governments. In order to keep up with the nations at the forefront of vaccine development, nations turned to state-sponsored cyberattacks in order to both hinder nations, and also obtain research and information about other countries' vaccine efforts. One such example was the threat group 'Cozy Bear', formally known as \acrfull{APT} 29. \acrshort{APT}29 used a number of tools to target various organisations involved in COVID-19 vaccine development in Canada, the United States and the United Kingdom. The \acrfull{NCSC} believe that the intention was highly likely stealing information and intellectual property relating to the vaccine \cite{APT29}.

@@

University attacks became a frequent headline in the UK as universities suffered attacks from different threat actors. A number of threat actors launched attacks against multiple universities in the hope to find a vulnerability in at least one. One such attack was aimed at both Newcastle University and Northumbria University, two universities in extremely close proximity \cite{newhack,norhack}. The attack crippled both Newcastle University and Northumbria University, however the attackers only managed to exfiltrate data from Newcastle University. Why was the attack successful on both occasions? Why wasn't knowledge of the attack shared?

@@

\begin{center}
	\begin{figure}[H]
		\includegraphics[width=\textwidth,keepaspectratio]{TanenbaumDistributed}
		\caption{A Distributed System visualised as \gls{middleware} \cite{tanenbaumdist}} 
		\label{fig:middleware}
	\end{figure}
\end{center}

@@

@@

\section{Aim} \label{aim}
The original aim for this project was to design and create a decentralised firewall that could communicate knowledge of cyberattacks aimed at universities in real-time, allowing other universities to protect themselves from the same attacks. This system would be distributed, and hence must conform to the previous description of a distributed system in section \ref{distributed}. Following an extensive amount of background reading, there appeared to be no existing implementation or design of such a system which inspired the alteration of the aim to instead focus on designing a protocol and implementing a \gls{stub} to demonstrate the protocol's effectiveness. This project will therefore not be implementing a firewall, but instead a system to coordinate firewalls. Further research determined that \gls{blockchain} was the best choice for the underlying structure for such a protocol, and so this final change shaped the current aim for this project: \textbf{Using Blockchain to Create a Decentralised Security Model for Distributed Systems}.

@@

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth,keepaspectratio]{rpc}
\caption{A breakdown of a \acrshort{RPC} \cite{tanenbaumdist}} 
\label{fig:rpc}
\end{figure}

@@

\section{Distributed Security}
The primary reference used for distributed security was Rivets and Lampson's \say{SDSI - A Simple Distributed Security Infrastructure} \cite{sdsi}. This publication provides an in-depth explanation of how a public-key infrastructure can be used in conjunction with access control lists to create a distributed security infrastructure. The majority of the literature within this publication is focused on creating and issuing certificates, something that is not relevant for this project, however Rivets and Lampson did provide clear requirements over the data structures within such a system. Rivets and Lampson implement a message system similar to that of Tanenbaum and van Steen's in section \ref{dist}. The message system proposed by Rivets and Lampson contains only a type and dictionary of attributes.

@@

\section{Fault Tolerance}
The primary reference for fault tolerance was Guerraoui and Schiper's \say{Fault-Tolerance by Replication in Distributed Systems}\cite{faulttol}. This literature details how replication can be used to provide fault tolerance in a distributed system in addition to ensuring consistency is maintained.  A number of backup techniques are discussed however the technique that is best suited for this project is primary backup replication. Primary backup replication consists of a client invoking an operation which is the applied to the primary data, and then cascaded to a number of additional backups.

@@

The type field holds an enum which is called MessageType and contains three values: advertisement, consensus and rule. Similarly the subtype field holds an enum that contains two values: broadcast and acknowledgement. Broadcast messages are used to advertise a new device, request consensus and distribute new a firewall transaction, whereas acknowledgements are used to provide a response, such as acknowledging the new device and providing consensus for a proposed transaction. As the system is decentralised, messages are not sent directly to remote hosts as there may be devices on the network that another device is unaware of, which would result in unsynchronised \gls{blockchain}s. The use of a hop count allows the number of times a message can be sent around the network to be limited, essential to prevent older messages from clogging up the system. The last three fields contain addresses of devices. It may appear unnecessary to provide these values as the messages are wrapped in an \acrshort{IP} packet which already contains these values. These are necessary as when messages are relayed, the \acrshort{IP} packet contains the address of the most recent sender, overwriting the original information. The source address contains the address of the origin device, the target address contains the address of the final destination, and the next address contains the address of the next device to propagate the message to.

\subsubsection{Advertisement Messages}
The advertisement message is used to advertise the presence of a new device within the distributed system. This message contains no additional fields than those described in table \ref{tab:common}. When the advertisement message subtype is broadcast, the source address is set to the new device's \acrshort{IP} address. At least one host must be known to advertise on the network, and this host's address will be placed into the next address field. The target address field is not used for broadcasts as there is no specific target that the message should be sent to. Once a node receives an advertisement broadcast, it will check whether that host is already known, if so, the message is ignored, and propagated, otherwise if the host is not yet known, the node first adds the new device to its list of known hosts, and then propagates the message. When the acknowledgement is sent, the target address is set to the address of the new device, which ensures that any other nodes which may have also advertised do not interpret the acknowledgement as regarding their broadcast. Once the acknowledgement arrives at the new device, the node adds the host which sent the ack, but does not propagate the message.

\subsubsection{Consensus Messages}
The consensus message is used when a client or firewall submits a new firewall rule to the \gls{stub} via \acrshort{IPC}. This message contains one additional field which holds the hash of the last block. When a node proposes a new firewall rule, it must first submit the hash of it's last block, which is sent to all known hosts. On receipt of a consensus broadcast, the node calculates the hash of the last block on it's chain, and if it matches the hash sent in the message it sends an ack, and appends the origin node's address to a list of pending rules. The node then forwards the message to all known hosts. After broadcasting the message, the origin host resets a counter, which is incremented every time it receives an ack from one of it's known hosts. If this counter reaches atleast half of it's known hosts in some bounded time, it is deemed to have obtained consensus and may now distribute the new firewall rule.

@@

\subsection{IPC}
There are four different types of \acrshort{IPC} message: rule, enable, disable, shutdown. The rule type is intended to be sent by a firewall, and contains a new firewall rule to be submitted to the distributed system. The enable and disable message types simply control the status of the stub, if it is believed that a host on the network may be malfunctioning (i.e. a firewall is sending bogus rules), it can be disabled and then re-enabled once the fault is fixed. Finally the shutdown message is used to properly terminate the \gls{stub}. Upon receiving the shutdown message, all sockets will be closed, the \acrshort{IPC} tunnel will be destroyed and the application will wait for all threads to terminate before exiting.
To ensure messages can be parsed correctly, all of these message types follow the same structure, and the rule field is simply disregarded for all non-rule messages.

@@

\subsection{Advertisement}
The advertisement message is sent at startup and is mutually exclusive from the other messages, as it is executed automatically and does not affect the consensus or rule transaction stages. The advertisement message is sent when the framework is executed. The framework will send an advertisement broadcast to all known hosts, each of which will relay the message to each of their known hosts, provided that the hop count has not yet been exceeded. Upon receiving an advertisement message, the host checks if the host is known, if not, it is added, and an acknowledgement is returned.

@@

\subsection{Consensus-Rule}
The consensus and rule messages are sent when a client or firewall sends an \acrshort{IPC} message to the local \gls{stub}. Upon receiving the rule, the \gls{stub} will marshal the parameters into a network message, which is then sent to all known hosts. When a node receives the consensus message, it will compare the hash in the message to the hash of the previous block on its chain, if the hashes match, the block is considered valid and an acknowledgement is returned. The broadcast is then forwarded to all of the hosts known by that node. When the origin node receives an acknowledgement, it will check if that host is known. If it is not, then it will not increment the ack count, as it would interfere with the consensus calculation, however it will still receive the rule, if consensus is achieved. The origin node will then wait for some timeout, after which it will test if a sufficient number of acknowledgements have been received in comparison to the number of known hosts. If this test passes (e.g. at least half of the known hosts must acknowledge), then a rule message is sent. No acknowledgements are sent upon receipt of a rule message, but nodes will continue to propagate the message until the hop limit is reached.

@@

\section{IPC}
In order to achieve cross-platform \acrshort{IPC}, different approaches were required for \acrfull{POSIX} compliant operating systems and Windows\cite{windows} operating systems. Each \acrshort{OS} type offers a number of \acrshort{IPC} methods, which are briefly discussed in the following sections.

The \gls{stub} provides the following abstractions for interfacing with native \acrshort{IPC}.\\\\

@@

\subsection{POSIX} \label{posixipc}
\acrshort{POSIX} systems provide four interfaces for \acrshort{IPC}: pipes, FIFO, message queues and sockets. Although pipes do allow processes to communicate, the primary purpose of pipes is to obtain the output of one command and pass this output into a second command, which is not suitable for our application. FIFO and message queues operate very similarly, whereby each method creates some file in the filesystem which can then be referenced by any process. FIFO is the predecessor to message queues, and hence different operating systems have varying implementations for FIFO, however as message queues are a \acrshort{POSIX} standard, all implementations provide the same interface, greatly simplifying the process of making this project cross compatible. For this reason, message queues were used for \acrshort{POSIX} \acrshort{IPC}. The final method is via UNIX sockets. The socket interface is primarily designed for networked communication, but also supports \acrshort{IPC}. Unfortunately, as sockets are intended for use in networks, some implementations still use the \acrshort{NIC} to communicate even locally, which slows down communication, and adds additional points of failure.

@@

\section{Sockets}
Network sockets are the primary method allowing processes to communicate over a network, however similarly to \acrshort{IPC}, \acrshort{POSIX} and Windows systems provide slightly different socket interfaces. A numeric port is bound to the socket which enables the network stack to communicate with the socket, and hence the process. The main difference is that Windows requires the initialisation of the WinSock\cite{winsock} library. Thankfully other than the initialisation of WinSock, the rest of the interface provided by WinSock is very similar to the basis of \acrshort{POSIX} sockets, which are known as Berkeley sockets. This similarity means that only small modifications need to be made to the source code. After creating the socket, it is then bound to the numeric port. The default ports used for this project are 8070 and 8071.

@@

\section{Multithreading}
In order for the \gls{stub} to be capable of consecutively sending and receiving messages over both the network, and via \acrshort{IPC}, the framework had to be multithreaded. Unsurprisingly, \acrshort{POSIX} and Windows use different thread models, which means there is no common threading interface. Thankfully, the \acrshort{POSIX} thread model has been ported to Windows in the form of the \acrshort{POSIX} Threads for Windows project\cite{pthread}. This allows the source code to be written identically for all operating systems, the only difference being that Windows will be linked with the necessary Dynamic-Link Library (DLL).

\section{Blockchain}
Once the previous \acrshort{API}s had been created, the \gls{blockchain} aspect of the project could be implemented. As the structure of these blocks have already been discussed in section \ref{firewallblock}, this section will focus on how the functionality of this was implemented, primarily the \gls{hashing} of blocks.

\subsection{Hash Algorithms} 
The first decision that had to be made was which hash algorithm to implement. A wide range of hash algorithms are available, however the most popular three are Secure Hash Algorithm (SHA), MD5 and BLAKE. The SHA hash suite boasts a number of algorithms which are classed into four generations: SHA-0, SHA-1, SHA-2 and SHA-3. The SHA-0 and SHA-1 algorithms are the least secure, and are now deemed insecure as collisions have been found, which would enable a malicious block to contain the hash of a valid block. At the time of writing this, there are no known collisions for the SHA-2 or SHA-3 families. The MD5 hash algorithm is also insecure for the same reason as SHA-1 and is hence not appropriate for this project. The final algorithm family is BLAKE, which consists of three families: BLAKE, BLAKE2 and BLAKE3. There are no known collisions within any of the BLAKE hash families, however the BLAKE algorithm has not been tested for \gls{blockchain} applications, whereas the SHA family has. For this reason, this project will utilise the same hash algorithm as Bitcoin \cite{bitcoin}, SHA-256.

\begin{table}[H]
\centering
\begin{tabular}{ |c|c|c|c|c| } 
\hline
\multicolumn{2}{|c|}{Hash Algorithm} & Output Size & Collisions found? & Used By \\
\hline
\multicolumn{2}{|c|}{MD5} & 128 & Yes & None \\
\hline
\multicolumn{2}{|c|}{SHA-0} & 160 & Yes & None \\
\hline
\multicolumn{2}{|c|}{SHA-1} & 160 & Yes & None \\
\hline
\multirow{6}*{SHA-2} & SHA-224 & 224 & No & None \\
& SHA-256 & 256 & No & Bitcoin\cite{bitcoin}, Bitcoin Cash \\
& SHA-384 & 384 & No & None \\
& SHA-512 & 512 & No & None \\
& SHA-512/224 & 224 & No & None \\
& SHA-512/256 & 256 & No & None \\
\hline
\multirow{4}*{SHA-3} & SHA3-224 & 224 & No & None \\
& SHA3-256 & 256 & No & None \\
& SHA3-384 & 384 & No & None \\
& SHA3-512 & 512 & No & None \\
\hline
\multirow{4}*{BLAKE} & BLAKE-224 & 224 & No & None \\
& BLAKE-256 & 256 & No & None \\
& BLAKE-384 & 384 & No & None \\
& BLAKE-512 & 512 & No & None \\
\hline
\multirow{2}*{BLAKE2} & BLAKE2s & 256 & No & Nano\cite{nano} \\
& BLAKE2b & 512 & No & None \\
\hline
\multicolumn{2}{|c|}{BLAKE3} & Variable & No & None \\
\hline
\end{tabular}
\caption{Comparison of Hash Algorithms}
\label{tab:hashalgo}
\end{table}

\subsection{Hash Implementation}
When implementing the desired hash algorithm, there were two primary options, to implement the algorithm from scratch, or to use an existing library. If the algorithm was to be implemented from scratch, then the resulting application would require less dependencies and would overall result in a more self-contained binary. Using a library would result in one additional dependency, however the vast number of algorithms provided by most libraries would allow the entire \gls{blockchain} and protocol to be easily updated should a more secure algorithm be preferable. Furthermore, most libraries are thoroughly tested and well engineered to guarantee the most efficient operation and make sure the algorithms can handle every edge case. For this reason it was decided that an external library would be implemented, and a breakdown of the considered libraries is provided below \cite{hashlibcomp}.

@@

From the previous table, it is apparent that libgcrypt supports the most algorithms from the comparison given in \cite{hashlibcomp}, however for cross compatibility, cryptlib boasts 37 supported operating systems. OpenSSL's libcrypto sits in the middle of the two, supporting nine hash algorithm families and 28 different operating systems. Additionally, libcrypto has been the default cryptography library distributed with most Linux, BSD and Windows operating systems, meaning it has received a large amount of support, and faced tight scrutiny by the likes of Microsoft. This made libcrypto the optimum library for use in this project.

\section{Fault Tolerance}
In regards to fault tolerance, this project was extremely limited, as it ran as a high-level application on general operating systems, meaning the operating systems themselves provided no guarantees of  performance or reliability. With regards to fault tolerance during operation, the only reasonable measure was to ensure no malformed data could be propagated throughout the system. Such malicious data could cause the local \gls{stub}, or even worse, remote \gls{stub}s to crash, rendering the entire system non-functional whilst the malicious data is removed from each system. This measure was achieved by using the fail-early principle in all functions within the system. When each function is executed, specific checks take place, such as ensuring buffers are large enough to store specified data, and that the data provided is valid, and of the required type. Many of these checks are provided by the standard library within the C language, such as verifying the string representation of an \acrshort{IP} address is  correctly formatted when parsing to a binary representation of the network address. The vast majority of these functions return an integer to represent whether the function exited normally, or if it exited with an error, which was replicated in many of the functions provided by this application. Returning the exit status enables any calling functions to know that the function did not exit normally, which means the output of that function will contain incorrect data, and so the path of execution can no longer follow the normal path, but must instead fail early, and return to the top-level calling function.

In addition to the fail-early principle, each node maintains a log of all complete transactions that have taken place on that node. Once a node receives either an advertisement, or a new rule, it is logged immediately. This means that should the operating system fail, or the system suffers from an unrecoverable fault, the operating system can immediately restart the system (which can run as a daemon) which will then load all known hosts into memory, and rebuild the chain from the transaction log.

\chapter{Testing} \label{test}
Testing the system required tests to be categorised into two types: local and network. Local tests consist of any functions which take place on the local node, such as \gls{hashing} of blocks and controlling network sockets. These local tests can be easily tested using unit tests. Network tests consist of protocol tests that can only be tested over a physical network. Network tests are much harder to strategically test an attempts to test these using unit tests would require careful synchronisation, such that when one node was testing it's send feature, another node would need to be ready to accept a message, for this reason networks tests did not use unit tests.
